Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO

### 2016-02-26 Paper/paper.tex
added discussion on multiple observations of identical sources

\usepackage[colorlinks, pdfborder={0 0 0}, plainpages=false]{hyperref}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage[usenames,dvipsnames]{color}
\usepackage{amssymb}
\usepackage[dvipsnames]{xcolor}

@@

% Macros for text notes and comments
\newcommand{\Note}[1]{\textcolor{blue}{\textbf{[#1]}}}
\newcommand{\lambdans}{\Lambda_\mathrm{NS}}
\newcommand{\chibh}{\chi_\mathrm{BH}}
\newcommand{\chins}{\chi_\mathrm{NS}}
\newcommand{\mbh}{m_\mathrm{BH}}
\newcommand{\mns}{m_\mathrm{NS}}
\newcommand{\mchirp}{\mathcal{M}_c}
\newcommand{\LL}{\mathcal{L}}

@@

\subsection{MCMC methods}
% % Describe the emcee-based PE code

@@

>>>>

\usepackage[colorlinks, pdfborder={0 0 0}, plainpages=false]{hyperref}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage[usenames,dvipsnames]{color}
\usepackage{amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{placeins}

@@

% Macros for text notes and comments
\newcommand{\Note}[1]{\textcolor{blue}{\textbf{[#1]}}}
\newcommand{\D}{\mathrm{d}}
\newcommand{\lambdans}{\Lambda_\mathrm{NS}}
\newcommand{\chibh}{\chi_\mathrm{BH}}
\newcommand{\chins}{\chi_\mathrm{NS}}
\newcommand{\mbh}{m_\mathrm{BH}}
\newcommand{\mns}{m_\mathrm{NS}}
\newcommand{\mchirp}{\mathcal{M}_c}
\newcommand{\LL}{\mathcal{L}}
\newcommand{\deff}{D_\mathrm{eff}}

@@

\subsection{Bayesian methods}
% % Describe the emcee-based PE code

@@

\pagebreak
\newpage
\FloatBarrier
\section{Combining observations: looking forward with Advanced LIGO}\label{s1:multiple_observations}
% 
% 
\subsection{Multiple identical sources at low SNR}\label{s2:identical_multiple}
% 
% 
Here we show a crude calculation which tells us that approximately $8$ observations
of NSBH signals with $\rho\geq 10$ distributed uniformly in effective volume, would
give us similarly accurate a measurement of $\lambdans$ as would a single detection
with $\rho=50$. We follow the outline in Ref.~\cite{Markakis:2010mp}.

We want to distribute signals in effective distance $\deff$, which is a combination of 
luminosity distance, inclination angle of the binary's orbital angular momentum
with respect to the detector, and its various sky angles. The signal strength 
scales inversely with $\deff$, if it comprises only of the dominant $(2,2)$ mode, allowing
for various angle-dependent factors to scale out beside the luminosity distance.
% 
In order to consider multiple events, lets imagine a population uniformly distributed
in effective volume, i.e. within a sphere of radius $\deff^\mathrm{max}$. The 
radius of this sphere is set by the lowest SNR that is distinguishable from noise
by LIGO searches. Now, dividie the sphere into $I$ shells of equal width. The radius
of the $i$-th sphere would then be $D_i = \deff^\mathrm{max} (i - 1/2)/I$. The 
measurement uncertainty in $\lambdans$ scales inversely with the SNR, and hence
directly with the effective distance to the source. Therefore, if we have a 
measurement error $\sigma_0$ for a source located at $\deff = D_0$, the same error
for the same source located within the $i$-th shell would be 
$$
\sigma_i = \sigma_0 \frac{D_i}{D_0}.
$$
Independent measurements of $\lambdans$ for identical sources at different distances
would have their combined error given by
\begin{equation}\label{eq:1oversigma}
\frac{1}{\sigma^2} = \sum_{i=1}^I \frac{N_i}{\sigma_i^2} = \left(\frac{D_0}{\sigma_0}\right)^2 \sum_{i=1}^I\frac{N_i}{D_i^2},
\end{equation}
where $N_i$ is the number of sources detected in the $i$-th shell, and is a random
variable, with its probability proportional to the volume of the shell, i.e.
$$
p(N_i) \propto \frac{V_i}{V_\mathrm{total}} \propto \dfrac{\left(\deff^\mathrm{max} \frac{i}{I}\right)^3 - \left(\deff^\mathrm{max} \frac{i-1}{I}\right)^3}{(\deff^\mathrm{max})^3} \propto \frac{1}{I^3} [i^3 - (i-1)^3].
$$
Therefore, the expected number of detections in the $i$-th shell would be
$$
\langle N_i\rangle = \frac{N}{I^3} [i^3 - (i-1)^3],
$$
where $N=\sum_{i=1}^I N_i$ is the total number of NSBH detections. $N$ is expected
to be Poisson distributed around the mean detection rate
$\mathcal{R}\equiv\langle N\rangle$. Let us assume that we have $\mathcal{R}_d$ detections
per year per unit effective volume, where $\mathcal{R}_d$ can vary from $0.6-1000$ per
$\mathrm{Gpc}^3$ per year~\cite{Abadie:2010cfa}. Then 
\begin{equation}
\langle N\rangle = \int_0^{\deff^\mathrm{max}} 4\pi D^2 \mathcal{R}_d \D D = \frac{4\pi}{3} \mathcal{R}_d (\deff^\mathrm{max})^3,
\end{equation}
and
\begin{eqnarray}
 \langle \frac{1}{\sigma^2}\rangle &=& \left(\frac{D_0}{\sigma_0}\right)^2 \int_0^{\deff^\mathrm{max}} \frac{4\pi\mathcal{R}_d D^2 }{D^2}\D D\\
 &=& \left(\frac{D_0}{\sigma_0}\right)^2 4\pi \mathcal{R}_d \deff^\mathrm{max},
\end{eqnarray}
where in the previous equation we have converted the summation in Eq.~\ref{eq:1oversigma} 
to an integral. So the root-mean-square averaged measurement error from $N$ sources 
distributed uniformly in effective volume within a sphere of radius $\deff^\mathrm{max}$
would be
\begin{equation}\label{eq:rmsSigmaIdenticalSources}
 \sigma_{avg} \equiv \frac{1}{\sqrt{\langle1/\sigma^{2}\rangle}} = \frac{\sigma_0}{D_0} \deff^\mathrm{max} \frac{1}{3\mathcal{R}_d}.
\end{equation}
Let us analyze the implications of Eq.~\ref{eq:rmsSigmaIdenticalSources}. It is
straightforward to deduce that the same measurement certainty as afforded by a single
observation with a high SNR $\alpha$ can be obtained from $N = \alpha^2/300$ observations
uniformly distributed in effective volume with SNRs $\geq 10$. E.g., to get to the level 
of certainty afforded at SNR$=70$, we would need $49/3\approx 16-17$ realistic detections.

There are two main caveats to this calculation, (i) it applies to identical sources, 
which is astrophysically next to impossible; (ii) the average measurement error 
$\sigma_{avg}$ has been defined as in the first equality in 
Eq.~\ref{eq:rmsSigmaIdenticalSources}.

% 
\subsection{Astrophysical source population}\label{s2:astro_multiple}
%
